Human Factors Influencing Visual Statistical Inference

### 2015-08-30 sociology_paper.Rnw
question: no sigma in model 1
figure 5

Similarly, we model the  time  an observer takes to identify a panel from a lineup. Let $Z_{ij}$ denote the logarithm of time taken for  observer $i$ to evaluate  lineup $j$. Let $\mu_{ij}=  E(Z_{ij})$ be the average of the (log) time taken by  observer $i$ to pick a panel from lineup $j$. We model this in a mixed effects model of the same structure as model~(\ref{eqn:demographic_response}) given as:
%\hh{
\begin{equation} \label{eqn:demographic_time}
Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
\end{equation}
where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model \eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ij} \sim N(0, \sigma^2)$.

@@

%\subsection{Model to Estimate Demographic Factor Effect} The feedback provided by each observer on a lineup is a binary response variable. Suppose $Y_{ij}$ denotes the response from observer $i$ on a  lineup $j$. $Y_{ij} =1$ if the response is correct otherwise $Y_{ij} =0$. Let $\pi_{ij}=  E(Y_{ij})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$. We model this in a generalized mixed effects model of the form
%\begin{equation} \label{eqn:demographic_response}
%g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
%\end{equation}
%where $\mu$ is an overall average probability for picking out the data plot from a lineup. $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender category $l(i)$, education level $m(i)$ and country name $s(i)$ respectively for observer $i$. $\ell_j$ is a random intercept predicting lineup difficulty level and we assume  $\ell_j \sim N(0, \sigma_\ell^2)$. $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
%
%
%The effect of demographic factors can be observed as time taken for each evaluation by an observer. Suppose $Z_{ij}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$. Let $\mu_{ij}=  E(Z_{ij})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$. We model this in a mixed effects model of the form
%%\hh{
%\begin{equation} \label{eqn:demographic_time}
%Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
%\end{equation}
%where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup. $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in Model \eqref{eqn:demographic_response}. $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.

@@

To explore the significance of the demographic factors model \eqref{eqn:demographic_response} is fit to the data, with age, country, education and gender as fixed effects. To estimate the significance of each of the factors, reduced models are fit with that factor removed from the model. Table~\ref{tbl:anova_factor} summarizes the results. Except for gender, all of the demographic factors show significant differences in  detection rates.  With respect to time taken, all of the demographic factors are significant.

@@

% latex table generated in R 3.2.2 by xtable 1.7-4 package
% Sat Aug 29 14:17:22 2015
\begin{table}[hbtp]
\caption{Parameter estimates of models~\eqref{eqn:demographic_time} and~\eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively. For time taken all the demographic factors are significant. For detection rate, age group 36-40, rest of the world and graduate degree are significantly different. For gender no difference in performance is observed. Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.}
\centering
\scalebox{.85}{
\begin{tabular}{rrr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
 && \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\

@@

\begin{table}[htbp]
\centering
\caption{Parameter estimates of models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively. For time taken all the demographic factors are significant. For detection rate, age group 36-40, rest of the world and graduate degree are significantly different. For gender no difference in performance is observed. Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.
%
}
\scalebox{0.86}{
\begin{tabular}{rlrrrrrrrrrr}
  \hline
Demographic&& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})} \\

\cline{3-6} \cline{8-11} 
Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
  \hline
\multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
&$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
\hline
Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
 &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.206 \\ 
  &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.038 \\ 
  &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
  &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.050 \\ 
  &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.143 \\ 
  &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.039 \\ 
  \hline
Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
  &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
  &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
  \hline
Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
 &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
  &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.331 \\ 
  &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.157 \\ 
  &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
  \hline
Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.133 \\ 
  \hline
\multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
& lineup($\sigma_{\ell}$) & 0.287 &  &  &  &   & 2.293 &  &  &  \\ 
 & Error($\sigma$) & 0.692 & &  &  &  & 0.893 && &\\ 
   \hline
\end{tabular}
}
\label{tbl:model_result_demographics}
\end{table}

Figure~\ref{fig:demographic_effect} supports the model results. The average time taken to respond (natural log) and detection rate are computed by lineup for the different demographic factor levels, and displayed using boxplots. Mean values are shown as dots inside the boxplot, as this statistic rather than the median is used in the models. %There are some small differences across demographic groups. The youngest age group (18-25) took least average time to evaluate a lineup. People from India took more time than others. People who have only a high school degree took less time and this is related to what we have seen for young age group. Averages of log time taken by male and female participants look similar but there are some differences as we see in Table~\ref{tbl:demographics}. 
The means and medians for response time (top row) are essentially equal. There are some differences between the two in the detection rate plots (bottom row). The differences in these values for different factors, that was tested by the models, can be seen in the plots. However, particularly for detection rate, the variation is huge. A large component of the variation is difficulty of the lineup. In some lineups the actual data plot was distinctively different from the null plots, and we would expect that the detection rate would be close to 100\%. In other lineups there was no difference, making it very difficult to evaluate, with low detection rate and longer time to respond. In Table~\ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}. This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.  So although the demographic factors emerge from the model to be statistically significant, the practical significance is minimal. We illustrate this with the following example of graduate degree.

@@

% This table had 4 digits $p$-value for last column
%\begin{table}[htbp]
%\centering
%\caption{Parameter estimates of Models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and probability of correct lineup evaluations respectively. For time taken all the demographic factors are significant. For probability of correct response age group 36-40, rest of the world and graduate degree are significantly different. For gender no difference in performance is observed. Lineup variability is estimated to be very large for Model \eqref{eqn:demographic_response}.}
%\scalebox{0.86}{
%\begin{tabular}{rlrrrrrrrrrr}
%  \hline
%Demographic&& \multicolumn{4}{c} {Model \eqref{eqn:demographic_time} Log Time} & &\multicolumn{4}{c} {Model \eqref{eqn:demographic_response} Proportion Correct} \\
%
%\cline{3-6} \cline{8-11} 
%Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
%  \hline
%\multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
%&$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
%\hline
%Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
% &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.2058 \\ 
%  &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.0376 \\ 
%  &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
%  &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.0501 \\ 
%  &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.1425 \\ 
%  &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.0394 \\ 
%  \hline
%Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
%  &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
%  &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
%  \hline
%Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
% &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
%  &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.3307 \\ 
%  &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.1566 \\ 
%  &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
%  \hline
%Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
%  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.1331 \\ 
%  \hline
%\multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
%& lineup($\sigma_{\ell}$) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
% & Error($\sigma$) & 0.479 & 0.692 &  &  & & && &\\ 
%   \hline
%\end{tabular}
%}
%\label{tbl:model_result_demographics}
%\end{table}

@@

%In Table~\ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}. This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.  Because of the huge impact of lineup variance, the practical impact of other factor variables on the probability of correct response is in fact very small. We illustrate this with the following example of graduate degree.

While some of the demographic factors are strongly statistically significant, the main source of variation in detection rate is the lineup difficulty. For example, let's examine the effect of graduate degree. To see just how large the effect is, we examine the change in detection rate for a (hypothetical) 18-25 year old female in the United States, with a graduate degree as compared to a high school degree, for an average difficulty lineup (random effect = 0). Plugging in the relevant quantities to the fitted model gives a difference equal to:
$$ \frac {\exp(-0.683+0.182)}{1+\exp(-0.683+0.182)}-  \frac {\exp(-0.683)}{1+\exp(-0.683)} = 0.377- 0.336 =0.041.$$ 
The person with a high school education on average picks the data plot in 33.5\% of lineups having average difficulty, as compared to 37.6\% if they have a graduate degree. This difference is reduced to 2\% for a lineup with one standard deviation order of magnitude difference in difficulty. For two standard deviations it further reduces to 0.3\%. Thus, although there is a statistically significant difference in the proportion at which participants identify the  data plot for some demographic factors, these are not  significant in any practical sense.  Figure~\ref{fig:practical_impact_graduate} illustrates this example showing fitted models for a US 18-25 female with either a high school education or a graduate degree.  Similar calculations show the same negligible impact of age level 36-40 (0.0533 at most)  and country (0.0424 at most) on the probability of a correct response. Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.

\begin{figure}[htbp] 
   \centering
%   \includegraphics[width=6.5in]{learning_trend.pdf} 
    \includegraphics[width=3.2in]{practical_impact_graduate.pdf} 
    \includegraphics[width=3.2in]{practical_impact_degree.pdf} 
   \caption{Proportion of estimated data identifications by an 18-25 year old female in the United States with a graduate degree compared to a high school degree. Even though having a graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.045 which is  negligible from a practical point of view. The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$) 
   of lineup variability.}
   \label{fig:practical_impact_graduate}
\end{figure}

@@

%Model \eqref{eqn:demographic_time} is fitted for the log time taken to evaluate a lineup and the results are also shown in  Table~\ref{tbl:model_result_demographics}.  The effect of all the demographic factors on log time taken to evaluate a lineup appears to be significant.

@@

\subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiments~5,~6, and~7 separately. In this model, attempt is fitted as a factor variable to allow for all possible non-linear learning trends.  It should be noted that we also examined an alternative to model \eqref{eqn:trend_response} where attempt was linearly fitted as a continuous covariate, but this also was not significant.

Table~\ref{tbl:model_result_response} displays the parameter estimates and $p$-values of fixed effect estimates of model \eqref{eqn:trend_response} examining detection rate. Only for experiment~6 there is some evidence of a learning curve. There are marginally small $p$-values for most of the attempt levels, and positive parameter estimates, suggesting that more attempts increases detection rate.

>>>>

Similarly, we model the  time  an observer takes to identify a panel from a lineup. Let $Z_{ij}$ denote the logarithm of time taken for  observer $i$ to evaluate  lineup $j$. Let $\mu_{ij}=  E(Z_{ij})$ be the average of the (log) time taken by  observer $i$ to pick a panel from lineup $j$. We model this in a mixed effects model of the same structure as model~(\ref{eqn:demographic_response}) given as:
%\hh{
\begin{equation} \label{eqn:demographic_time}
Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
\end{equation}
where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model~\eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ij} \sim N(0, \sigma^2)$.

@@

%\subsection{Model to Estimate Demographic Factor Effect} The feedback provided by each observer on a lineup is a binary response variable. Suppose $Y_{ij}$ denotes the response from observer $i$ on a  lineup $j$. $Y_{ij} =1$ if the response is correct otherwise $Y_{ij} =0$. Let $\pi_{ij}=  E(Y_{ij})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$. We model this in a generalized mixed effects model of the form
%\begin{equation} \label{eqn:demographic_response}
%g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
%\end{equation}
%where $\mu$ is an overall average probability for picking out the data plot from a lineup. $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender category $l(i)$, education level $m(i)$ and country name $s(i)$ respectively for observer $i$. $\ell_j$ is a random intercept predicting lineup difficulty level and we assume  $\ell_j \sim N(0, \sigma_\ell^2)$. $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
%
%
%The effect of demographic factors can be observed as time taken for each evaluation by an observer. Suppose $Z_{ij}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$. Let $\mu_{ij}=  E(Z_{ij})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$. We model this in a mixed effects model of the form
%%\hh{
%\begin{equation} \label{eqn:demographic_time}
%Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
%\end{equation}
%where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup. $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in Model~\eqref{eqn:demographic_response}. $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.

@@

To explore the significance of the demographic factors model~\eqref{eqn:demographic_response} is fit to the data, with age, country, education and gender as fixed effects. To estimate the significance of each of the factors, reduced models are fit with that factor removed from the model. Table~\ref{tbl:anova_factor} summarizes the results. Except for gender, all of the demographic factors show significant differences in  detection rates.  With respect to time taken, all of the demographic factors are significant.

@@

% latex table generated in R 3.2.2 by xtable 1.7-4 package
% Sat Aug 29 14:17:22 2015
\begin{table}[hbtp]
\caption{\label{tbl:model_result_demographics}Parameter estimates of models~\eqref{eqn:demographic_time} and~\eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively. For time taken all the demographic factors are significant. For detection rate, age group 36-40, rest of the world and graduate degree are significantly different. For gender no difference in performance is observed. Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.}
\centering
\scalebox{.85}{
\begin{tabular}{rrr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
 && \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\

@@

% \begin{table}[htbp]
% \centering
% \caption{Parameter estimates of models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively. For time taken all the demographic factors are significant. For detection rate, age group 36-40, rest of the world and graduate degree are significantly different. For gender no difference in performance is observed. Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.
% %
% }
% \scalebox{0.86}{
% \begin{tabular}{rlrrrrrrrrrr}
%   \hline
% Demographic&& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})} \\
% 
% \cline{3-6} \cline{8-11} 
% Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
%   \hline
% \multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
% &$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
% \hline
% Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
%  &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.206 \\ 
%   &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.038 \\ 
%   &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
%   &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.050 \\ 
%   &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.143 \\ 
%   &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.039 \\ 
%   \hline
% Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
%   &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
%   &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
%   \hline
% Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
%  &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
%   &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.331 \\ 
%   &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.157 \\ 
%   &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
%   \hline
% Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
%   &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.133 \\ 
%   \hline
% \multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
% & lineup($\sigma_{\ell}$) & 0.287 &  &  &  &   & 2.293 &  &  &  \\ 
%  & Error($\sigma$) & 0.692 & &  &  &  & 0.893 && &\\ 
%    \hline
% \end{tabular}
% }
% \label{tbl:model_result_demographics}
% \end{table}

Figure~\ref{fig:demographic_effect} supports the model results. The average time taken to respond (natural log) and detection rate are computed by lineup for the different demographic factor levels, and displayed using boxplots. Mean values are shown as dots inside the boxplot, as this statistic rather than the median is used in the models. %There are some small differences across demographic groups. The youngest age group (18-25) took least average time to evaluate a lineup. People from India took more time than others. People who have only a high school degree took less time and this is related to what we have seen for young age group. Averages of log time taken by male and female participants look similar but there are some differences as we see in Table~\ref{tbl:demographics}. 
The means and medians for response time (top row) are essentially equal. There are some differences between the two in the detection rate plots (bottom row). The differences in these values for different factors, that was tested by the models, can be seen in the plots. However, particularly for detection rate, the variation is huge. A large component of the variation is difficulty of the lineup. In some lineups the actual data plot was distinctively different from the null plots, and we would expect that the detection rate would be close to 100\%. In other lineups there was no difference, making it very difficult to evaluate, with low detection rate and longer time to respond. In Table~\ref{tbl:model_result_demographics} notice that lineup specific error is estimated as 2.29 which is much higher than any of the other parameter estimates in the model. This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.  So although the demographic factors emerge from the model to be statistically significant, the practical significance is minimal. We illustrate this with the following example of graduate degree.

@@

% This table had 4 digits $p$-value for last column
%\begin{table}[htbp]
%\centering
%\caption{Parameter estimates of Models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and probability of correct lineup evaluations respectively. For time taken all the demographic factors are significant. For probability of correct response age group 36-40, rest of the world and graduate degree are significantly different. For gender no difference in performance is observed. Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.}
%\scalebox{0.86}{
%\begin{tabular}{rlrrrrrrrrrr}
%  \hline
%Demographic&& \multicolumn{4}{c} {Model~\eqref{eqn:demographic_time} Log Time} & &\multicolumn{4}{c} {Model~\eqref{eqn:demographic_response} Proportion Correct} \\
%
%\cline{3-6} \cline{8-11} 
%Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
%  \hline
%\multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
%&$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
%\hline
%Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
% &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.2058 \\ 
%  &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.0376 \\ 
%  &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
%  &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.0501 \\ 
%  &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.1425 \\ 
%  &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.0394 \\ 
%  \hline
%Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
%  &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
%  &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
%  \hline
%Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
% &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
%  &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.3307 \\ 
%  &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.1566 \\ 
%  &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
%  \hline
%Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
%  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.1331 \\ 
%  \hline
%\multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
%& lineup($\sigma_{\ell}$) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
% & Error($\sigma$) & 0.479 & 0.692 &  &  & & && &\\ 
%   \hline
%\end{tabular}
%}
%\label{tbl:model_result_demographics}
%\end{table}

@@

%In Table~\ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model~\eqref{eqn:demographic_response}. This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.  Because of the huge impact of lineup variance, the practical impact of other factor variables on the probability of correct response is in fact very small. We illustrate this with the following example of graduate degree.

While some of the demographic factors are strongly statistically significant, the main source of variation in detection rate is the lineup difficulty. For example, let's examine the effect of graduate degree. To see just how large the effect is, we examine the change in detection rate for a (hypothetical) 18-25 year old female in the United States, with some graduate course work as compared to an undergraduate degree, for an average difficulty lineup (random effect of zero). Plugging in the relevant quantities to the fitted model gives a difference equal to:
$$ \frac {\exp(-0.61-0.17)}{1+\exp(-0.61-0.17)}-  \frac {\exp(-0.61+0.12)}{1+\exp(-0.61+0.12)} = 0.314- 0.379 =-0.065.$$ 
The person with some graduate course work picks the data plot  on average in 33.5\% of lineups of average difficulty, as compared to 37.9\% if they have an undergraduate degree. This difference is reduced to 2\% for a lineup with one standard deviation order of magnitude difference in difficulty. For two standard deviations it further reduces to 0.3\%. Thus, although there is a statistically significant difference in the proportion at which participants identify the  data plot for some demographic factors, these are not  significant in any practical sense.  Figure~\ref{fig:practical_impact_graduate} illustrates this example showing fitted models for a US 18-25 female with either a high school education or a graduate degree.  Similar calculations show the same negligible impact of age level 36-40 (\Sexpr{round(exp(-0.61+0.31)/(1+ exp(-0.61+0.31)) - exp(-0.61)/(1 + exp(-0.61)),3)} at most)  and country (\Sexpr{round(exp(-0.61+0.08)/(1+ exp(-0.61+0.08)) - exp(-0.61-0.18)/(1 + exp(-0.61-0.18)), 3)} at most) on the probability of a correct response. Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.

\begin{figure}[htbp] 
   \centering
%   \includegraphics[width=6.5in]{learning_trend.pdf} 
    \includegraphics[width=3.2in]{practical_impact_graduate.pdf} 
    \includegraphics[width=3.2in]{practical_impact_degree.pdf} 
   \caption{Proportion of estimated data identifications by an 18-25 year old female in the United States with a graduate degree compared to a high school degree. Even though having a graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.045 which is  negligible from a practical point of view. The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.29$) 
   of lineup variability.}
   \label{fig:practical_impact_graduate}
\end{figure}

@@

%Model~\eqref{eqn:demographic_time} is fitted for the log time taken to evaluate a lineup and the results are also shown in  Table~\ref{tbl:model_result_demographics}.  The effect of all the demographic factors on log time taken to evaluate a lineup appears to be significant.

@@

\subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiments~5,~6, and~7 separately. In this model, attempt is fitted as a factor variable to allow for all possible non-linear learning trends.  It should be noted that we also examined an alternative to model~\eqref{eqn:trend_response} where attempt was linearly fitted as a continuous covariate, but this also was not significant.

Table~\ref{tbl:model_result_response} displays the parameter estimates and $p$-values of fixed effect estimates of model~\eqref{eqn:trend_response} examining detection rate. Only for experiment~6 there is some evidence of a learning curve. There are marginally small $p$-values for most of the attempt levels, and positive parameter estimates, suggesting that more attempts increases detection rate.
